---
title: "Covid91 vaccine study Final2023"
author: "Bella Verdin"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "Spring 2024" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))

male = filter(FinalData, sex =="M")
female = filter(FinalData, sex =="F")
lgbtq = filter(FinalData, LGBTQ =="gay")
drugUser = filter(FinalData, DrugUser =="yes")

```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

$H_0$ : There are NO statistically significant difference in the data, meaning the vaccine and placebo are equally as effective.

$H_A$ : There are statistically significant difference in the data, meaning the vaccine and placebo are NOT equally as effective.

#Methods

I will create a different data set for each sub population. then I will use my helper file cat~cat to create a bar chart to display the data.  then after I will use xtabs to get the percentages for the data. Since only two types are being tested, there is no need for any loop to measure multiple variables. 

## Males

### Graphical Results 

```{r}
barchartGC(~infected + treatment,data=male)
barchartGC(~infected + treatment,data=male, type="percent")
```

Based on the graph, there was a difference of 30% in covid-19 and a 70% with the placebo within the people who received the vaccination and the ones who got Covid-19. There was a slight difference with those who got the the vaccination an were "ok."

### Numerical Results

```{r}
tableMale <- xtabs(~infected + treatment, data=male)
rowPerc(tableMale)
colPerc(tableMale)
```

The table shows that there was x2 time difference with those who received the vaccine and still got covid-19. 70% of the ones who took the placebo did not get the virus,compared to the 30$ who got the vaccine. There was no noticeable distinction in those who got the vaccine and were okay. Its effective on males in this group. 

### Inferential Results

```{r}
chisq.test(tableMale)
chisqtestGC(tableMale)
maleFisher <- fisher.test(tableMale)
maleFisher
```

The p-value for males is 0.0000000000001098. The p-value is small which is why I can not reject the null hypothesis. the vaccine is not effective in preventing Covid-19 among the males. Based of the Fisher's Exact Test the p-value is 0.00000000000004631 showing a difference in getting covid-19 and being a male. If you're a male, the chances of getting covid-19 are the same if you take the placebo or get the vaccine. 


## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions