---
title: "Covid91 vaccine study Final2023"
author: "Bella Verdin"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "Spring 2024" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))

male = filter(FinalData, sex =="M")
female = filter(FinalData, sex =="F")
lgbtq = filter(FinalData, LGBTQ =="gay")
drugUser = filter(FinalData, DrugUser =="yes")

```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

$H_0$ : There are NO statistically significant difference in the data, meaning the vaccine and placebo are equally as effective.

$H_A$ : There are statistically significant difference in the data, meaning the vaccine and placebo are NOT equally as effective.

#Methods

I will create a different data set for each sub population. then I will use my helper file cat~cat to create a bar chart to display the data.  then after I will use xtabs to get the percentages for the data. Since only two types are being tested, there is no need for any loop to measure multiple variables. 

## Males

### Graphical Results 

```{r}
barchartGC(~infected + treatment,data=male)
barchartGC(~infected + treatment,data=male, type="percent")
```

Based on the graph, there was a difference of 30% in covid-19 and a 70% with the placebo within the people who received the vaccination and the ones who got Covid-19. There was a slight difference with those who got the the vaccination an were "ok."

### Numerical Results

```{r}
tableMale <- xtabs(~infected + treatment, data=male)
rowPerc(tableMale)
colPerc(tableMale)
```

The table shows that there was x2 time difference with those who received the vaccine and still got covid-19. 70% of the ones who took the placebo did not get the virus,compared to the 30$ who got the vaccine. There was no noticeable distinction in those who got the vaccine and were okay. Its effective on males in this group. 

### Inferential Results

```{r}
chisq.test(tableMale)
chisqtestGC(tableMale)
maleFisher <- fisher.test(tableMale)
maleFisher
```

The p-value for males is 0.0000000000001098. The p-value is small which is why I can not reject the null hypothesis. the vaccine is not effective in preventing Covid-19 among the males. Based of the Fisher's Exact Test the p-value is 0.00000000000004631 showing a difference in getting covid-19 and being a male. If you're a male, the chances of getting covid-19 are the same if you take the placebo or get the vaccine. 


## Females

### Graphical results 

```{r}
barchartGC(~infected + treatment,data=female)
barchartGC(~infected + treatment,data=female, type="percent")
```

According to the graph, those who received covid-19 showed a difference between 40% and the placebo showing a 60% in those who took the vaccine. The Females who received the vaccine adn were OK didn't show any noticeable difference. 

### Numerical Results 

```{r}
tableFemale <- xtabs(~infected + treatment, data=female)
rowPerc(tableFemale)
colPerc(tableFemale)
```

The table shows that there is a 20% difference in the vaccine rates with those who receive the coivd-19 vaccine. And the females who received the vaccine was 40% and 60% for the placebo. there is no difference with those received he vaccine. 

### Inferential Results

```{r}
chisq.test(tableFemale)
chisqtestGC(tableFemale)
femaleFisher <- fisher.test(tableFemale)
femaleFisher
```

The Chi-squared p-value is 0.002732 and the p-value is small, meaning the vaccine is not effective in preventing covid-19 in the females. Based on the Fisher"s Exact test the p-value is 0.0002162. I fail to reject the null hypothesis, due to the p-value being too small. 


## LGBTQ

###Graphical Results

```{r}
barchartGC(~infected + treatment,data=lgbtq)
barchartGC(~infected + treatment,data=lgbtq, type="percent")
```

Based on the graph, the individuals who received the vaccine and had covid-19 ad a 80% compared to a 20% placebo difference. There wasn't a apparent variance with those who received the vaccine and were OK. 

### Numerical Results 

```{r}
tableLGBTQ <- xtabs(~infected + treatment, data=lgbtq)
rowPerc(tableLGBTQ)
colPerc(tableLGBTQ)
```

On the table is shows a 40% difference in individuals who got covid-19 and got the vaccine. It shows about 80% of those who received the placebo had a 20% chance of not getting sick. There wasn't a discernible difference with those who received the vaccine and were OK. 

###Inferential Results 

```{r}
chisq.test(tableLGBTQ)
chisqtestGC(tableLGBTQ)
lgbtqFisher <- fisher.test(tableLGBTQ)
lgbtqFisher
```

The Chi-Squared value is 0.00000006668 and its too small. So that shows the vaccine is not effective in preventing the spread of covid-19 in the LGBTQ. With the Fisher's Exact test the p-value is 0.00000001029. Meaning I fail to reject the null hypothesis. by taking the vaccine or placebo you will still have the same odds in getting covid-19.

## Druggies

###Graphical Results 

```{r}
barchartGC(~infected + treatment,data=drugUser)
barchartGC(~infected + treatment,data=drugUser, type="percent")
```

The graph shows those who had covid-19. The druggies who did get the vaccine showed no difference between the 88% with covid-19and about a 10% chance with the placebo.

###Numercial Results

```{r}
tableDrug <- xtabs(~infected + treatment, data=drugUser)
rowPerc(tableDrug)
colPerc(tableDrug)
```

The table shows that the druggies who took the vaccine and got covid-19 showed a difference. The druggies that took the placebo had a 12% chance of not being sick and the druggies had a 88%. those that were OK and took the vaccine showed no significant difference. 

###Inferential Results 

```{r}
chisq.test(tableDrug)
chisqtestGC(tableDrug)
drugFisher <- fisher.test(tableDrug)
drugFisher
```

The Chi-squared p-value is 0.000000002842 is small. so the vaccine is not effective in preventing covid-19 from spreading among other people. With the Fisher's exact test the p-value is 0.000000000232. Meaning I fail to reject the null hypothesis. With all the tests ran is shows that the vaccine does not prevent covid-19. 


# Overall Results and Conclusions

I wanted to figure out is Covid-19 was prevented in the male group. I came to the conclusion that the vaccine is ineffective in protecting males against covid-19. I fail to reject the null hypothesis for the male group. I also fail to reject the null hypothesis as the vaccine was also ineffective in the female group. the results of the research shows that the vaccination did not succeed in preventing covid-19, and with LGBTQ in the tests as well, I fail to reject the null hypothesis once again. Sam with the drug users as well there was no vaccine prevention on covid-19 so I reject the null hypothesis.  



